Hypercoercions and a Framework for \\ Equivalence of Cast Calculi

### 2019-12-26 paper.tex
fix more typos
revised, following suggestions from the WGT reviewers

@@

Around 2006, several groups of researchers proposed ways to integrate
dynamic typing and static typing, notably gradual typing
\citep{siek2006gradual}, hybrid typing \citep{flanagan2006hybrid},
migratory typing \citep{Tobin-Hochstadt:2006fk}, and multi-language
interoperability \citep{Gray:2005ij,Matthews:2007zr}. Researchers
usually define the semantics of gradually typed languages by
translation to an intermediate language with casts, such as the blame
calculus \citep{wadler2009well} and other cast calculi
\citep{siek2009exploring}. Unfortunately, straightforward
implementations of casts on higher-order values (functions, objects,
etc.) impose significant runtime overheads that can change the
asymptotic space complexity of a program
\citep{herman2010space}. There are several known space-efficient cast
representations, with various strengths and weaknesses
\citep{siek2015blame,siek2010threesomes,garcia2013calculating,kuhlenschmidt2018efficient,siek2012interpretations,garcia2014deriving}.
The current state of the art includes

\begin{itemize}
\item threesomes \citep{siek2010threesomes,garcia2013calculating},
\item supercoercions \citep{garcia2013calculating}, and
\item coercions in normal form
  \citep{siek2012interpretations,siek2015blame}.
\end{itemize}
%Recall that in these systems, casts are compressed using a composition
%operator.
These systems compress casts using a compose operator.
Threesomes and supercoercions are good for mechanized
metatheory because their compose operators are structurally recursive,
making them easy to define in a proof assistant such as Agda. In
contrast, the coercions in normal form have compose operators that are
not structurally recursive, which makes it more difficult to define in
Agda, requiring what amounts to an explicit proof of termination.
%
On the other hand, coercions in normal form are easier to understand
than threesomes (with a strange labeled bottom type), and
supercoercions (10 different kinds).

This paper presents a new cast representation, named
\emph{hypercoercions}, that is good for both mechanized metatheory and
good for implementation. The composition operator for hypercoercions
is defined by structural recursion and hypercoercions are suggestive
of a bit-level representation that minimizes the need for pointers and
fits all first-order casts into 64 bits.
%
We present two flavors of hypercoercions to support the two blame
tracking strategies from the literature: D and
UD~\citep{siek2009exploring}. We are interested in the D blame
tracking strategy because it comes with a more straightforward notion of
safe cast compared to UD~\citep{siek2009exploring}, which is why D was
chosen in Grift compiler~\citep{kuhlenschmidt2018efficient}. We are also 
interested in UD because it plays a
prominent role in the gradual typing literature
\citep{wadler2009well}.  The UD hypercoercions were inspired by the
supercoercions of \citet{garcia2013calculating} (hence the name) and
the D hypercoercions were inspired by the normal forms of
\citet{siek2012interpretations}.
%
The semantics of casts can be lazy or
eager~\citep{siek2009exploring}. In this paper, we focus on lazy cast
strategies because we suspect that they are more efficient than eager
strategies and because \citet{new2019gradual} show that the eager
strategies are incompatible with $\eta$-equivalence of functions.

@@

\begin{figure}
  Syntax
  \[
  \begin{array}{lrcl}
  \stxrule{T, S}{Types}{
    \star \mid{}
    P
  }
  \stxrule{P}{Pre-types}{
    \POOunit \mid
    \POOfun{T}{T} \mid
    \POOprod{T}{T} \mid
    \POOsum{T}{T}
  }
  \stxrule{e}{Terms}{
    \eOOvar{x} \mid{}
    \eOOsole{} \mid{}
    \eOOlam{T}{T}{x}{e} \mid
    \eOOapp{e}{e} \mid
    \eOOcons{T}{T}{e}{e} \mid
    \eOOcar{e} \mid
    \eOOcdr{e}
  }
  \stxrulecont{
    \eOOinl{T}{T}{e} \mid
    \eOOinl{T}{T}{e} \mid
    \eOOcase{e}{e}{e} \mid
    \eOOcast{e}{T}{l}{T} \mid
    \eOOblame{l}
  }
  \end{array}
  \]

@@

The syntax and static semantics are the same for the \lazyD{} and
\lazyUD{} cast calculi. They are reviewed in \figref{fig:blame-static}.
As usual, the important features are the cast expressions,
$\eOOcast{e}{T_1}{l}{T_2}$, which are responsible for runtime type
checking, and blame expressions, $\eOOblame{l}$, that raise errors.
The syntax and static semantics is the same that of
\citet{siek2009exploring} except for a few minor exceptions:
\begin{itemize}
\item We add sum, product, and unit types.
\item We separate types into those with a type constructor at the top,
  the \emph{pretypes} ($\POOunit$, functions, products, and sums),
  versus the dynamic type $\star$ (a.k.a. $\mathtt{Dyn}$ or
  $\mathbb{?}$).
\item We annotate the type of many expressions explicitly because we 
refer to them in the dynamic semantics. These expressions include cons 
expressions, left injections, right injections, and lambda abstractions.
\end{itemize}
As usual, the source $T_1$ and target types $T_2$ of a cast
$\eOOcast{e}{T_1}{l}{T_2}$ are required to be consistent, written $T_1
\sim T_2$. The consistency relation is standard and defined in
\figref{fig:blame-static}.

@@

\redrule{
\sOOreturn{v_2}{\kOOappII{(\vOOfun{\mathcal{E}}{x}{e})}{\kappa}}}{
\sOOinspect{e}{\mathcal{E}[x:=v_2]}{\kappa}}{}
  \hiredrule{
    \sOOreturn{v_1}{
      \kOOappII{\vOOcast{v_2}{
          \cOOcast{\POOfun{T_1}{T_2}}{l}{\POOfun{T_3}{T_4}}
      }}{\kappa}
%     (\mathtt{app_2} \; \vOOcast{v_2}{
%       \cOOcast{\POOfun{T_1}{T_2}}{l}{\POOfun{T_3}{T_4}}
%     } \; \kappa)
    }
  }{
    \sOOreturn{v_1}{
      \kOOcast{\cOOcast{T_3}{l}{T_1}}{
        \kOOappII{v_2}{
          \kOOcast{\cOOcast{T_2}{l}{T_4}}{\kappa}
        }
      }
%     \langle\rangle
%     (\mathtt{app_2} \; v_2 \; 
%     \langle\cOOcast{T_2}{l}{T_4}\rangle \kappa}
    }
  }{}
  \redrule{
  \sOOreturn{
    \vOOcons{v_1}{v_2}
  }{\kOOcar{\kappa}}
  }{
  \sOOreturn{v_1}{\kappa}
  }{}
  \hiredrule{
    \sOOreturn{
      \vOOcast{v}{\cOOcast{\POOprod{T_1}{T_2}}{l}{
          \POOprod{T_3}{T_4}}}
    }{\mathtt{fst} \; \kappa}
  }{
    \sOOreturn{v}{
      \kOOcar{
        \kOOcast{\cOOcast{T_1}{l}{T_3}}{\kappa}
      }}
  }{}
  \redrule{
  \sOOreturn{
  \vOOcons{v_1}{v_2}
  }{\kOOcdr{\kappa}}
  }{
  \sOOreturn{v_2}{\kappa}
  }{}

  \hiredrule{
    \sOOreturn{
      \vOOcast{v}{\cOOcast{\POOprod{T_1}{T_2}}{l}{
          \POOprod{T_3}{T_4}}}
    }{\mathtt{snd} \; \kappa}
  }{
    \sOOreturn{v}{
      \kOOcdr{\kOOcast{\cOOcast{T_2}{l}{T_4}}{\kappa}}}
  }{}

@@

\redrule{
\sOOreturn{v_3}{
  \kOOcaseIII{\vOOinl{v}}{v_2}{\kappa}
}
}{
\sOOreturn{v}{\kOOappII{v_2}{\kappa}}
}{}

\redrule{
  \sOOreturn{v_3}{
    \kOOcaseIII{\vOOinr{v}}{v_2}{\kappa}
  }
}{
\sOOreturn{v}{\kOOappII{v_3}{\kappa}}
}{}

\redrule{
  \highlight{\sOOreturn{v_3}{
    \kOOcaseIII{
        \vOOcast{v}{\cOOcast{\POOsum{T_1}{T_2}}{l}{\POOsum{T_3}{T_4}}}
      }{v_2}{\kappa}
  }}
}{
  \highlight{\sOOreturn{v_3'}{
      \kOOcaseIII{v}{v_2'}{\kappa}
  }}
}{\\&&
\highlight{\text{where}\;
\kappa\;\text{is waiting for a value of type}\;T
}
\\&&
\highlight{\text{and}\;
	v2' = \vOOcast{v_2}{\cOOcast{\POOfun{T_4}{T}}{l}{\POOfun{T_2}{T}}}}
\\&&
\highlight{\text{and}\;
v3' = \vOOcast{v_3}{\cOOcast{\POOfun{T_3}{T}}{l}{\POOfun{T_1}{T}}}}
}

@@

In this section, we review the
coercions~\citep{henglein1994dynamic,herman2010space} and the normal form of
\citet{siek2012interpretations} to motivate the design of hypercoercions.  We
omit sum types and product types in this section because
\citet{siek2012interpretations} did not discuss them. We assume a basic
familiarity with coercions, and suggest that readers unfamiliar with coercions
to familiarize themselves with \citet{siek2012interpretations} and
\citet{siek2015blame}.

@@

  \comprule{
    \hyperCoercionC{h_1}{m_1}{t_1}
  }{
    \hyperCoercionC{?^l}{m_2}{t_2}
  }{
    \hyperCoercionC{h_1}{m_1}{\bot^l}
  }{
    \sidecond{
       m_1 \not\smile m_2
      \; \text{and} \;
      \forall l. t \neq \bot^{l}
    }
  }

\comprulel{
\hyperCoercionC{h_1}{m_1}{t_1}
}{
\hyperCoercionC{\epsilon}{m_2}{t_2}
}{
\hyperCoercionC{h_1}{m_1}{\bot^l}
}{
\sidecond{
  m_1 \not\smile m_2
  \; \text{and} \;
  \forall l. t \neq \bot^{l}
}
}
\comprule{
\hyperCoercionC{h_1}{m_1}{t_1}
}{
\hyperCoercionC{?^l}{m_2}{t_2}
}{
\hyperCoercionC{h_1}{m_1 \fatsemi^{l} m_2}{t_2}
}{
\sidecond{
  m_1 \smile m_2
  \; \text{and} \;
  \forall l. t \neq \bot^{l}
}
}
  \comprule{
    \hyperCoercionC{h_1}{m_1}{t_1}
  }{
    \hyperCoercionC{\epsilon}{m_2}{t_2}
  }{
    \hyperCoercionC{h_1}{m_1 \fatsemi^{\ell} m_2}{t_2}
  }{
    \sidecond{
      m_1 \smile m_2
      \; \text{and} \;
      \forall l. t \neq \bot^{l}
    }
  }
  \end{array}
  \]

@@

\begin{definition}[Lazy D Cast ADT]
  \label{def:surely-lazyd}
  A  Cast is a Lazy D Cast if:
  \begin{enumerate}
  \item If $v : T$ then $applyCast(v,id(T)) = \mathtt{succ} \; v $
  \item If $\judgeType{v}{T_1}$ and
    $ \judgeTypeFT{c_1}{T_1}{T_2}$ and
    $ \judgeTypeFT{c_2}{T_2}{T_3}$\\
    then $applyCast(v,seq(c_1,c_2)) = 
    applyCast(v,c_1) >>= \lambda v'.applyCast(v',c_2)$\\
    where 
    \[
    \begin{array}{rcl}
      \rOOsucc{v} >>= f & = & f(v) \\
      \rOOfail{l} >>= f & = & \rOOfail{l}
    \end{array}
    \]
  \item If $v : T_1$ and $T_1 \not\smile T_2$,
    then $applyCast(v,cast(T_1, l, T_2)) = \rOOfail{l} $
  \item If $v : \star$,
    then $applyCast(v,cast(\TOOdyn,l,\TOOdyn)) = \rOOsucc{v} $
  \item If $v : P$,
    then $applyCast(\hcvOOinj{P}{v},cast(\star,l,P')) 
    = applyCast(v,cast(P,l,P')) $
  \item If $v : P$,
    then $applyCast(v,cast(P,l,\star)) = \rOOsucc{(\hcvOOinj{P}{v})} $
  \item If $v : \POOunit$,
    then $applyCast(v,cast(\POOunit,l,\POOunit)) = \rOOsucc{v} $
  \item If $(\hcvOOfun{c_1}{\mathcal{E}}{x}{e}{c_2}) : \POOfun{T_1}{T_2}$,
    then\\
    $ 
    applyCast(\hcvOOfun{c_1}{\mathcal{E}}{x}{e}{c_2}, 
    cast(\POOfun{T_1}{T_2},l,\POOfun{T_3}{T_4})) \\
    = 
    \rOOsucc{(\hcvOOfun{seq(cast(T_3,l,T_1),c_1)}{\mathcal{E}}{x}{b}{seq(c_2,cast(T_2,l,T_4))})}$
  \item If $(\hcvOOcons{v_1}{c_1}{v_2}{c_2}) : \POOprod{T_1}{T_2}$,
    then \\
    $ 
    applyCast(\hcvOOcons{v_1}{c_1}{v_2}{c_2},cast(\POOprod{T_1}{T_2},l,T_3 
    \times 
    T_4))$\\
    $ = 
    \rOOsucc{(\hcvOOcons{v_1}{seq(c_1,cast(T_1,l,T_3))}{v_2}{seq(c_2,cast(T_2,l,T_4))})}
    $ 
  \item If $(\hcvOOinl{v}{c}) : \POOsum{T_1}{T_2}$,
    then \\
    $ 
    applyCast(\hcvOOinl{v}{c},cast(\POOsum{T_1}{T_2},l,\POOsum{T_3}{T_4}))
    = \rOOsucc{(\hcvOOinl{v}{seq(c,cast(T_1,l,T_3))})} $
  \item If $(\hcvOOinr{v}{c}) : \POOsum{T_1}{T_2}$,
    then \\$
    applyCast(\hcvOOinr{v}{c},cast(\POOsum{T_1}{T_2},l,\POOsum{T_3}{T_4}))
    = \rOOsucc{(\hcvOOinr{v}{seq(c,cast(T_2,l,T_4))})} $
  \end{enumerate}
\end{definition}

@@

  Transition \fbox{$\judgeSreduce{C}{s}{s}$}
  \[
  \begin{array}{rclr}
  \hiredruleS{
    \sOOinspect{\eOOcast{e}{T_1}{l}{T_2}}{\mathcal{E}}{\kappa}
  }{
    \sOOinspect{e}{\mathcal{E}}{ext(cast(T_1,l,T_2),\kappa)}
  }{}
  \redruleS{
    \sOOinspect{\eOOvar{x}}{\mathcal{E}}{\kappa}
  }{  
    \sOOreturn{\mathcal{E}(x)}{\kappa}
  }{}
  \redruleS{
    \sOOinspect{\eOOsole}{\mathcal{E}}{\kappa}
  }{
    \sOOreturn{\hcvOOtt}{\kappa}
  }{}
  \hiredruleS{
    \sOOinspect{\eOOlam{T_1}{T_2}{x}{e}}{\mathcal{E}}{\kappa}
  }{
    \sOOreturn{\hcvOOfun{id(T_1)}{\mathcal{E}}{x}{e}{id(T_2)}}{\kappa}
  }{}
  \redruleS{
    \sOOinspect{\eOOcons{T_1}{T_2}{e_1}{e_2}}{\mathcal{E}}{\kappa}
  }{
    \sOOinspect{e_1}{\mathcal{E}}{
      wrap(\kOOconsI{T_1}{T_2}{e_2}{\mathcal{E}}{\kappa})
    }
  }{}
  \redruleS{
    \sOOinspect{\eOOinl{T_1}{T_2}{e}}{\mathcal{E}}{\kappa}
  }{
    \sOOinspect{e}{\mathcal{E}}{wrap(\kOOinl{T_1}{T_2}{\kappa})}
  }{}
  \redruleS{
    \sOOinspect{\eOOinl{T_1}{T_2}{e}}{\mathcal{E}}{\kappa}
  }{
    \sOOinspect{e}{\mathcal{E}}{wrap(\kOOinr{T_1}{T_2}{\kappa})}
  }{}
  \redruleS{
    \sOOinspect{\eOOapp{e_1}{e_2}}{\mathcal{E}}{\kappa}
  }{
    \sOOinspect{e_1}{\mathcal{E}}{wrap(\kOOappI{e_2}{\mathcal{E}}{\kappa})}}{}

@@

  Apply continuation
  \fbox{$\continue{k}{v} = s$}
  \[
  \begin{array}{rclr}

@@

\hifunrule{
  \continue{\hcvOOcons{v_1}{c_1}{v_2}{c_2}}{
    \kOOcar{\kappa}
  }
}{
  \sOOreturn{v_1}{ext(c_1,k)}
}{}
\hifunrule{
  \continue{\hcvOOcons{v_1}{c_1}{v_2}{c_2}}{
    \kOOcdr{\kappa}
  }
}{
  \sOOreturn{v_2}{ext(c_2,k)}
}{}

@@

\fbox{$applyCast(v,c) = r$}
\[
\begin{array}{rclr}
\funrule{
  applyCast(v,\lcOOnull)
}{
  \rOOsucc{v}
}{}
\funrule{
  applyCast(v,\lcOOcons{s}{c})
}{
  applyStep(v,s) >>= \lambda v. applyCast(v,c)
}{}
\end{array}
\]

>>>>

\newcommand{\mbind}{\ensuremath{\underline{\gg}}}

@@

Around 2006, several groups of researchers proposed ways to integrate
dynamic typing and static typing, notably gradual typing
\citep{siek2006gradual}, hybrid typing \citep{knowles2010hybrid},
migratory typing \citep{Tobin-Hochstadt:2006fk}, and multi-language
interoperability \citep{Gray:2005ij,Matthews:2007zr}. Researchers
usually define the semantics of gradually typed languages by
translation to an intermediate language with casts, such as the blame
calculus \citep{wadler2009well} and other cast calculi
\citep{siek2009exploring}. Unfortunately, straightforward
implementations of casts on higher-order values (functions, objects,
etc.) impose significant runtime overheads that can change the
asymptotic space complexity of a program
\citep{herman2010space}. There are several known space-efficient cast
representations, with various strengths and weaknesses
\citep{Siek:2015:BCT:2737924.2737968,Siek:2010:TWB:1706299.1706342,Garcia:2013:CTB:2500365.2500603,kuhlenschmidt2018efficient,siek2012interpretations,garcia2014deriving}.
The current state of the art includes

\begin{itemize}
\item threesomes 
\citep{Siek:2010:TWB:1706299.1706342,Garcia:2013:CTB:2500365.2500603},
\item supercoercions \citep{Garcia:2013:CTB:2500365.2500603}, and
\item coercions in normal form
  \citep{siek2012interpretations,Siek:2015:BCT:2737924.2737968}.
\end{itemize}
%Recall that in these systems, casts are compressed using a composition
%operator.
These systems compress casts using a compose operator.
Threesomes and supercoercions are good for mechanized
metatheory because their compose operators are structurally recursive,
making them easy to define in a proof assistant such as Agda. In
contrast, the coercions in normal form have compose operators that are
not structurally recursive, which makes it more difficult to define in
Agda, requiring what amounts to an explicit proof of termination.
%
On the other hand, coercions in normal form are easier to understand
than threesomes (with a strange labeled bottom type), and
supercoercions (10 different kinds).

This paper presents a new cast representation, named
\emph{hypercoercions}, that is good for both mechanized metatheory and
good for implementation. The composition operator for hypercoercions
is defined by structural recursion and hypercoercions are suggestive
of a bit-level representation that minimizes the need for pointers and
fits all first-order casts into 64 bits.
%
We present two flavors of hypercoercions to support the two blame
tracking strategies from the literature: D and
UD~\citep{siek2009exploring}. We are interested in the D blame
tracking strategy because it comes with a more straightforward notion of
safe cast compared to UD~\citep{siek2009exploring}, which is why D was
chosen in the Grift compiler~\citep{kuhlenschmidt2018efficient}. We are also 
interested in UD because it plays a
prominent role in the gradual typing literature
\citep{wadler2009well}.  The UD hypercoercions were inspired by the
supercoercions of \citet{Garcia:2013:CTB:2500365.2500603} (hence the name) and
the D hypercoercions were inspired by the normal forms of
\citet{siek2012interpretations}.
%
The semantics of casts can be lazy or
eager~\citep{siek2009exploring}. In this paper, we focus on lazy cast
strategies because we suspect that they are more efficient than eager
strategies and because \citet{new2019gradual} show that the eager
strategies are incompatible with $\eta$-equivalence of functions.

@@

\begin{figure}
  Syntax
  \[
  \begin{array}{lrcl}
  \stxrule{T, S}{Types}{
    \star \mid{}
    P
  }
  \stxrule{P}{Pre-types}{
    \POOunit \mid
    \POOfun{T}{T} \mid
    \POOprod{T}{T} \mid
    \POOsum{T}{T}
  }
  \stxrule{e}{Terms}{
    \eOOvar{x} \mid{}
    \eOOsole{} \mid{}
    \eOOlam{T}{T}{x}{e} \mid
    \eOOapp{e}{e} \mid
    \eOOcons{T}{T}{e}{e} \mid
    \eOOcar{e} \mid
    \eOOcdr{e}
  }
  \stxrulecont{
    \eOOinl{T}{T}{e} \mid
    \eOOinr{T}{T}{e} \mid
    \eOOcase{e}{e}{e} \mid
    \eOOcast{e}{T}{l}{T} \mid
    \eOOblame{l}
  }
  \end{array}
  \]

@@

The syntax and static semantics are the same for the \lazyD{} and
\lazyUD{} cast calculi. They are reviewed in \figref{fig:blame-static}.
As usual, the important features are the cast expressions,
$\eOOcast{e}{T_1}{l}{T_2}$, which are responsible for runtime type
checking, and blame expressions, $\eOOblame{l}$, that raise errors. Polarities 
of blame labels is not treated but are straightforward to incorporate. 
The syntax and static semantics is the same as that of
\citet{siek2009exploring} except for a few minor exceptions:
\begin{itemize}
\item We add sum, product, and unit types.
\item We separate types into those with a type constructor at the top,
  the \emph{pretypes} ($\POOunit$, functions, products, and sums),
  versus the dynamic type $\star$ (a.k.a. $\mathtt{Dyn}$ or
  $\mathbb{?}$).
\item We annotate the type of many expressions explicitly because we 
refer to them in the dynamic semantics. These expressions include cons 
expressions, left injections, right injections, and lambda abstractions.
\end{itemize}
As usual, the source $T_1$ and target types $T_2$ of a cast
$\eOOcast{e}{T_1}{l}{T_2}$ are required to be consistent, written $T_1
\sim T_2$. The consistency relation is standard and defined in
\figref{fig:blame-static}.

@@

\redrule{
\sOOreturn{v_2}{\kOOappII{(\vOOfun{\mathcal{E}}{x}{e})}{\kappa}}}{
\sOOinspect{e}{\mathcal{E}[x:=v_2]}{\kappa}}{}
  \hiredrule{
    \sOOreturn{v_1}{
      \kOOappII{\vOOcast{v_2}{
          \cOOcast{\POOfun{T_1}{T_2}}{l}{\POOfun{T_3}{T_4}}
      }}{\kappa}
%     (\mathtt{app_2} \; \vOOcast{v_2}{
%       \cOOcast{\POOfun{T_1}{T_2}}{l}{\POOfun{T_3}{T_4}}
%     } \; \kappa)
    }
  }{
    \sOOreturn{v_1}{
      \kOOcast{\cOOcast{T_3}{l}{T_1}}{
        \kOOappII{v_2}{
          \kOOcast{\cOOcast{T_2}{l}{T_4}}{\kappa}
        }
      }
%     \langle\rangle
%     (\mathtt{app_2} \; v_2 \; 
%     \langle\cOOcast{T_2}{l}{T_4}\rangle \kappa}
    }
  }{}
  \redrule{
  \sOOreturn{
    \vOOcons{v_1}{v_2}
  }{\kOOcar{\kappa}}
  }{
  \sOOreturn{v_1}{\kappa}
  }{}
  \hiredrule{
    \sOOreturn{
      \vOOcast{v}{\cOOcast{\POOprod{T_1}{T_2}}{l}{
          \POOprod{T_3}{T_4}}}
    }{\kOOcar{\kappa}}
  }{
    \sOOreturn{v}{
      \kOOcar{
        \kOOcast{\cOOcast{T_1}{l}{T_3}}{\kappa}
      }}
  }{}
  \redrule{
  \sOOreturn{
  \vOOcons{v_1}{v_2}
  }{\kOOcdr{\kappa}}
  }{
  \sOOreturn{v_2}{\kappa}
  }{}

  \hiredrule{
    \sOOreturn{
      \vOOcast{v}{\cOOcast{\POOprod{T_1}{T_2}}{l}{
          \POOprod{T_3}{T_4}}}
    }{\kOOcdr{\kappa}}
  }{
    \sOOreturn{v}{
      \kOOcdr{\kOOcast{\cOOcast{T_2}{l}{T_4}}{\kappa}}}
  }{}

@@

\redrule{
\sOOreturn{v_3}{
  \kOOcaseIII{(\vOOinl{v})}{v_2}{\kappa}
}
}{
\sOOreturn{v}{\kOOappII{v_2}{\kappa}}
}{}

\redrule{
  \sOOreturn{v_3}{
    \kOOcaseIII{(\vOOinr{v})}{v_2}{\kappa}
  }
}{
\sOOreturn{v}{\kOOappII{v_3}{\kappa}}
}{}

\redrule{
  \highlight{\sOOreturn{v_3}{
    \kOOcaseIII{
        \vOOcast{v}{\cOOcast{\POOsum{T_1}{T_2}}{l}{\POOsum{T_3}{T_4}}}
      }{v_2}{\kappa}
  }}
}{
  \highlight{\sOOreturn{v_3'}{
      \kOOcaseIII{v}{v_2'}{\kappa}
  }}
}{\\&&
\highlight{\text{where}\;
	\kappa:T
%\kappa\;\text{is waiting for a value of type}\;T
}
\\&&
\highlight{\text{and}\;
v2' = \vOOcast{v_2}{\cOOcast{\POOfun{T_3}{T}}{l}{\POOfun{T_1}{T}}}}
\\&&
\highlight{\text{and}\;
v3' = \vOOcast{v_3}{\cOOcast{\POOfun{T_4}{T}}{l}{\POOfun{T_2}{T}}}}
}

@@

In this section, we review the
coercions~\citep{henglein1994dynamic,herman2010space} and the normal form of
\citet{siek2012interpretations} to motivate the design of hypercoercions.  We
omit sum types and product types in this section because
\citet{siek2012interpretations} did not discuss them. We assume a basic
familiarity with coercions, and suggest that readers unfamiliar with coercions
to familiarize themselves with \citet{siek2012interpretations} and
\citet{Siek:2015:BCT:2737924.2737968}.

@@

  \comprule{
    \hyperCoercionC{h_1}{m_1}{t_1}
  }{
    \hyperCoercionC{?^l}{m_2}{t_2}
  }{
    \hyperCoercionC{h_1}{m_1}{\bot^l}
  }{
    \sidecond{
       m_1 \not\smile m_2
      \; \text{and} \;
      \forall l. t_1 \neq \bot^{l}
    }
  }

\comprulel{
\hyperCoercionC{h_1}{m_1}{t_1}
}{
\hyperCoercionC{\epsilon}{m_2}{t_2}
}{
\hyperCoercionC{h_1}{m_1}{\bot^l}
}{
\sidecond{
  m_1 \not\smile m_2
  \; \text{and} \;
  \forall l. t_1 \neq \bot^{l}
}
}
\comprule{
\hyperCoercionC{h_1}{m_1}{t_1}
}{
\hyperCoercionC{?^l}{m_2}{t_2}
}{
\hyperCoercionC{h_1}{m_1 \fatsemi^{l} m_2}{t_2}
}{
\sidecond{
  m_1 \smile m_2
  \; \text{and} \;
  \forall l. t_1 \neq \bot^{l}
}
}
  \comprule{
    \hyperCoercionC{h_1}{m_1}{t_1}
  }{
    \hyperCoercionC{\epsilon}{m_2}{t_2}
  }{
    \hyperCoercionC{h_1}{m_1 \fatsemi^{\ell} m_2}{t_2}
  }{
    \sidecond{
      m_1 \smile m_2
      \; \text{and} \;
      \forall l. t_1 \neq \bot^{l}
    }
  }
  \end{array}
  \]

@@

\begin{definition}[Lazy D Cast ADT]
  \label{def:surely-lazyd}
  A  Cast is a Lazy D Cast if:
  \begin{enumerate}
  \item If $v : T$ then $applyCast(v,id(T)) = \mathtt{succ} \; v $
  \item If $\judgeType{v}{T_1}$ and
    $ \judgeTypeFT{c_1}{T_1}{T_2}$ and
    $ \judgeTypeFT{c_2}{T_2}{T_3}$\\
    then $applyCast(v,seq(c_1,c_2)) = 
    applyCast(v,c_1) \mbind{} \lambda v'.applyCast(v',c_2)$\\
    where 
    \[
    \begin{array}{rcl}
      \rOOsucc{v} \mbind{} f & = & f(v) \\
      \rOOfail{l} \mbind{} f & = & \rOOfail{l}
    \end{array}
    \]
  \item If $v : T_1$ and $T_1 \not\smile T_2$,
    then $applyCast(v,cast(T_1, l, T_2)) = \rOOfail{l} $
  \item If $v : \star$,
    then $applyCast(v,cast(\TOOdyn,l,\TOOdyn)) = \rOOsucc{v} $
  \item If $v : P$,
    then $applyCast(\hcvOOinj{P}{v},cast(\star,l,P')) 
    = applyCast(v,cast(P,l,P')) $
  \item If $v : P$,
    then $applyCast(v,cast(P,l,\star)) = \rOOsucc{(\hcvOOinj{P}{v})} $
  \item If $v : \POOunit$,
    then $applyCast(v,cast(\POOunit,l,\POOunit)) = \rOOsucc{v} $
  \item If $(\hcvOOfun{c_1}{\mathcal{E}}{x}{e}{c_2}) : \POOfun{T_1}{T_2}$,
    then\\
    $ 
    applyCast(\hcvOOfun{c_1}{\mathcal{E}}{x}{e}{c_2}, 
    cast(\POOfun{T_1}{T_2},l,\POOfun{T_3}{T_4})) \\
    = 
    \rOOsucc{(\hcvOOfun{seq(cast(T_3,l,T_1),c_1)}{\mathcal{E}}{x}{e}{seq(c_2,cast(T_2,l,T_4))})}$
  \item If $(\hcvOOcons{v_1}{c_1}{v_2}{c_2}) : \POOprod{T_1}{T_2}$,
    then \\
    $ 
    applyCast(\hcvOOcons{v_1}{c_1}{v_2}{c_2},cast(\POOprod{T_1}{T_2},l,T_3 
    \times 
    T_4))$\\
    $ = 
    \rOOsucc{(\hcvOOcons{v_1}{seq(c_1,cast(T_1,l,T_3))}{v_2}{seq(c_2,cast(T_2,l,T_4))})}
    $ 
  \item If $(\hcvOOinl{v}{c}) : \POOsum{T_1}{T_2}$,
    then \\
    $ 
    applyCast(\hcvOOinl{v}{c},cast(\POOsum{T_1}{T_2},l,\POOsum{T_3}{T_4}))
    = \rOOsucc{(\hcvOOinl{v}{seq(c,cast(T_1,l,T_3))})} $
  \item If $(\hcvOOinr{v}{c}) : \POOsum{T_1}{T_2}$,
    then \\$
    applyCast(\hcvOOinr{v}{c},cast(\POOsum{T_1}{T_2},l,\POOsum{T_3}{T_4}))
    = \rOOsucc{(\hcvOOinr{v}{seq(c,cast(T_2,l,T_4))})} $
  \end{enumerate}
\end{definition}

@@

  Transition \fbox{$\judgeSreduce{C}{s}{s}$}
  \[
  \begin{array}{rclr}
  \hiredruleS{
    \sOOinspect{\eOOcast{e}{T_1}{l}{T_2}}{\mathcal{E}}{\kappa}
  }{
    \sOOinspect{e}{\mathcal{E}}{ext(cast(T_1,l,T_2),\kappa)}
  }{}
  \redruleS{
    \sOOinspect{\eOOvar{x}}{\mathcal{E}}{\kappa}
  }{  
    \sOOreturn{\mathcal{E}(x)}{\kappa}
  }{}
  \redruleS{
    \sOOinspect{\eOOsole}{\mathcal{E}}{\kappa}
  }{
    \sOOreturn{\hcvOOtt}{\kappa}
  }{}
  \hiredruleS{
    \sOOinspect{\eOOlam{T_1}{T_2}{x}{e}}{\mathcal{E}}{\kappa}
  }{
    \sOOreturn{\hcvOOfun{id(T_1)}{\mathcal{E}}{x}{e}{id(T_2)}}{\kappa}
  }{}
  \redruleS{
    \sOOinspect{\eOOcons{T_1}{T_2}{e_1}{e_2}}{\mathcal{E}}{\kappa}
  }{
    \sOOinspect{e_1}{\mathcal{E}}{
      wrap(\kOOconsI{T_1}{T_2}{e_2}{\mathcal{E}}{\kappa})
    }
  }{}
  \redruleS{
    \sOOinspect{\eOOinl{T_1}{T_2}{e}}{\mathcal{E}}{\kappa}
  }{
    \sOOinspect{e}{\mathcal{E}}{wrap(\kOOinl{T_1}{T_2}{\kappa})}
  }{}
  \redruleS{
    \sOOinspect{\eOOinr{T_1}{T_2}{e}}{\mathcal{E}}{\kappa}
  }{
    \sOOinspect{e}{\mathcal{E}}{wrap(\kOOinr{T_1}{T_2}{\kappa})}
  }{}
  \redruleS{
    \sOOinspect{\eOOapp{e_1}{e_2}}{\mathcal{E}}{\kappa}
  }{
    \sOOinspect{e_1}{\mathcal{E}}{wrap(\kOOappI{e_2}{\mathcal{E}}{\kappa})}}{}

@@

  Apply continuation
  \fbox{$\continue{v}{k} = s$}
  \[
  \begin{array}{rclr}

@@

\hifunrule{
  \continue{\hcvOOcons{v_1}{c_1}{v_2}{c_2}}{
    \kOOcar{\kappa}
  }
}{
  \sOOreturn{v_1}{ext(c_1,\kappa)}
}{}
\hifunrule{
  \continue{\hcvOOcons{v_1}{c_1}{v_2}{c_2}}{
    \kOOcdr{\kappa}
  }
}{
  \sOOreturn{v_2}{ext(c_2,\kappa)}
}{}

@@

\fbox{$applyCast(v,c) = r$}
\[
\begin{array}{rclr}
\funrule{
  applyCast(v,\lcOOnull)
}{
  \rOOsucc{v}
}{}
\funrule{
  applyCast(v,\lcOOcons{s}{c})
}{
  applyStep(v,s) \mbind{} \lambda v. applyCast(v,c)
}{}
\end{array}
\]
